Detecting Spam on Social Networking Sites

### 2012-05-08 docs/paper/paper.tex
update

We decided to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports. In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy. We first extracted a number of fields that we expected might improve classification.

@@

\textbf{Message Body \& Subject:} We extracted both the message

\textbf{Sender IP:} The IP address of the sender at the time that the user sent the message.

@@

\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message.

@@

@@

\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups. 48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%. In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.

figure \ref{fig:usaspam}
figure \ref{fig:usahamfig:lambda}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
    \label{fig:usaspam}
\end{figure}

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:drop}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

>>>>

We chose to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports. In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy. We first extracted a number of fields that we expected might improve classification.

@@

\textbf{Message Body \& Subject:} We extracted both the message body and subject from 
each spam and ham message. The cleaning script generated two additional features based 
on this. First, we counted the frequency of non-standard punctuation (we noticed that many 
spam messages would, for example, put spaces before commas, periods and quotation marks, 
and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
lowercase letters in the text, after observing relatively high amounts of uppercase text in spam messages.
%% add numbers

@@

\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message. We saw 
that the mean reply rate for ham messages was 81.77\% (SD: 38.61), while it was only 0.81\% (SD: 8.97) for spam messages.

@@

\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as 
revealed by their IP address differed remarkably between the spam and ham user groups. 
48\% of spam users claimed to be from the United States, with the next most popular 
countries being the UK, Ghana, Senegal, Germany and Canada taking combined 25\%. 
In contrast, only 15\% ham users claimed to be from the United States, with Russia, 
Korea, UK, France and Germany combining for 24\%. This distribution of profile countries can 
be seen in Figure \ref{fig:cprof}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

Users are free to choose any country on their profile. We record each IP address from which a 
user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP 
database) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
countries for spammers are significantly different, while the top ham countries are 
virtually unchanged. Furthermore, only 30\% of the IP addresses of spam users who 
claimed to be from the United States actually mapped back to a US-based ISP, with a combined 
46\% indicating a Ghana, Nigeria and Senegal ISP (Figure \ref{fig:usaspam}). This contrasts 
with the 73\% that US-based IP addresses comprised for ham users whose profiles stated that 
they were in the United States (Figure \ref{fig:usaham}).

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
    \label{fig:usaspam}
\end{figure}

@@

@@

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:drop}
figure \ref{fig:friends}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:roc-25}
